### 2021-09-30 paper/header.tex
paper

% **************GENERATED FILE, DO NOT EDIT**************

\title{Applied Measure Theory for Probabilistic Modeling}

\author[1]{Chad Scherrer}
\author[2]{Moritz Schauer}
\affil[1]{Informative Prior}
\affil[2]{Chalmers and Gothenburg University, Sweden}

\keywords{Julia, Measure theory, Probability, Statistics}

\hypersetup{
pdftitle = {Applied Measure Theory for Probabilistic Modeling},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Chad Scherrer, Moritz Schauer},
pdfkeywords = {Julia, Measure theory, Probability, Statistics},
}